\documentclass{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{fullpage}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{enumitem}
\usepackage{url}


\parindent0in
\pagestyle{plain}
\thispagestyle{plain}

\newcommand{\assignment}{Homework 1}
\newcommand{\duedate}{July 19, 2022}

% \renewcommand\thesubsection{\arabic{subsection}}

\title{Homework 1}
\date{}

\begin{document}

Fundação Getulio Vargas\hfill\\
Estruturas de Dados e Algoritmos\hfill\textbf{\assignment}\\
Prof.\ Jorge Poco\hfill\textbf{Due:}: \duedate\\
\smallskip\hrule\bigskip

{\let\newpage\relax\maketitle}
\maketitle


\section{Induction [3pts]}
Answers should be written in this document. 

\begin{enumerate}
  \item Prove by Induction that:
  \( \sum_{i=1}^{n}i^2=\frac{n(n+1)(2n+1)}{6} \qquad\forall n \geq 0\)

  \item Prove by Induction that:
  $\forall n \geq 7$ it is true $3^n<n!$
  
  \item Prove by Induction that $\forall n \geq 0$
  \[
    \left \lceil\frac{n}{2} \right \rceil=
    \left\{
    \begin{array}{ll}
    \frac{n}{2}& \textrm{si $n$ es par}\\
    \frac{n+1}{2}& \textrm{si $n$ es impar}
    \end{array}
    \right.
  \]
  
    \item Prove by induction that a number is divisible by 3 if and only if the sum of its digits is divisible by 3.
    
    \item Prove that any integer greater than 59 can be formed using only 7 and 11 cent coins.
  
  \item Prove by induction that $F_{n+k}=F_{k}F_{n+1}+F_{k-1}F_{n}$
  
  \item Prove by induction in $n$ that \(\sum_{m=0}^{n}{n \choose m}=2^n\)
  
  \item Prove by induction that a graph with $n$ vertices can have at most  $\frac{n(n-1)}{2}$ edges.
  
  \item Prove by induction that a full binary tree\footnote{\url{http://web.cecs.pdx.edu/~sheard/course/Cs163/Doc/FullvsComplete.html}} with $n$ levels has $2^n-1$ vertices.
  
  \item A polygon is convex if each pair of points in the polygon can be joined by a straight line that does not leave the polygon. Prove by induction in $n>3$ that the sum of the angles of a polygon of $n$ vertices is $180(n-2)$.
  
\end{enumerate}

\section{Correctness of bubblesort [3pts]}
Bubblesort is a popular, but inefficient, sorting algorithm. It works by repeatedly swapping adjacent elements that are out of order.

\begin{algorithm}[H]
\SetAlgoLined
  \For{$i = 1$ \textbf{to} $A.length -1$} {
    \For{$j = A.length$ \textbf{downto} $i + 1$} {
      \If{$A[j] < A[j-1]$} {
        exchange $A[j]$ with $A[j-1]$
      }
    }
  }
\caption{BUBBLESORT(A)}
\end{algorithm}

\begin{enumerate}[label=\Alph*]
  \item \textbf{(0.5pt)} Let $A'$ denote the output of BUBBLESORT(A). To prove that BUBBLESORT is correct, we need to prove that it terminates and that

  \begin{equation} \label{eq:1}
    A'[1] \leq A'[2] \leq ... \leq A'[n]
  \end{equation}

  where $n = A.length$. In order to show that BUBBLESORT actually sorts, what else do we need to prove?

  The next two parts will prove inequality~(\ref{eq:1}).

  \item \textbf{(1.0pt)} State precisely a loop invariant \textbf{for} the for loop in lines 2–6, and prove that this loop invariant holds. Your proof should use the structure of the loop invariant proof presented in this chapter.

  \item \textbf{(1.0pt)} Using the termination condition of the loop invariant proved in part (B), state a loop invariant for the for loop in lines 1–7 that will allow you to prove inequality~(\ref{eq:1}). Your proof should use the structure of the loop invariant proof presented in this chapter.

  \item \textbf{(0.5pt)} What is the worst-case running time of BUBBLESORT? How does it compare to the running time of insertion sort?
\end{enumerate}


\section{Insertion Sort - Mergesort - Quicksort[4pts]}
Implement the insertion sort, mergesort and quicksort using the template \texttt{test.py} (use Python 3.X). Your code should print the tuple (number of objects, time insertion\_sort, time merge\_sort, time quick\_sort)

You must submit \texttt{test.py}. Graphs and descriptions must be included in this document. 

\subsection{Random Order}
\begin{enumerate}
  \item Create 10 sets of numbers in random order. The sets must have \{10k, 20k, 30k, ..., 100k\} numbers.
  
  \item Sort these numbers using the 3 algorithms and calculate the time each algorithm takes for each set of numbers.
  
  \item Generate a plot (using excel or another tool) showing a \emph{line chart}, where the $x$-axis is the ``number of elements", and the $y$-axis is the time that the algorithms took in Python. This plot must have 3 lines of different colors with a legend.
  
  \item Write a small paragraph (3 to 4 lines) describing the results.
\end{enumerate}

\subsection{Ascending Order}
Do the same experiment when the numbers are ordered in ascending order.

\subsection{Descending Order}
Do the same experiment when the numbers are ordered in descending order.



\end{document}